---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #2

```{r}

library(remotes)
remotes::install_github("heike/classdata")
library(classdata)

```


# Summary of Finding: 


# Question 1: Inspect the first few lines of the data set

```{r}

?ames 
colnames(ames)
head(ames)
str(ames)
summary(ames$AC)

```

**Parcel ID**:\
Format: Character\
Description: Unique identifier for each property\
Expected Data Range: Unique IDs, length of characters varies\

**Address**:\
Format: Character\
Description: Property address in Ames, IA\
Expected Data Range: Various addresses in Ames, IA\

**Style**:\
Format: Factor\
Description: Type of housing, such as "Colonial," "Ranch," "Split-Level," etc.\
Expected Data Range: Different housing styles as factor with 12 levels\

**Occupancy**:\
Format: Factor\
Description: Type of occupancy, indicating whether the property is occupied or not\
Expected Data Range: Occupancy status as factor with 5 levels\

**Sale Date**:\
Format: Date\
Description: Date of the property sale\
Expected Data Range: Dates ranging from 2017-07-03 to 2022-08-31\

**Sale Price**:\
Format: Numeric (in US dollars)\
Description: Sales price of the property\
Expected Data Range: Sale prices in US dollars, varying from 0 to 20500000\

**Multi Sale**:\
Format: Character (logical- boolean)\
Description: Indicates if the sale was part of a package or multiple sale\
Expected Data Range: Y/N\

**YearBuilt**:\
Format: Numeric (integer)\
Description: Year in which the house was built\
Expected Data Range: Years ranging from 0 to 2022\

**Acres**:\
Format: Numeric\
Description: Size of the lot in acres\
Expected Data Range: Acres of the lots, ranging from 0 to 12.0120\

**TotalLivingArea (sf)**:\
Format: Numeric (Square Feet)\
Description: Total living area of the property in square feet\
Expected Data Range: Living area sizes in square feet, varying from 0 to 6007\

**Bedrooms**:\
Format: Numeric\
Description: Number of bedrooms in the property\
Expected Data Range: Number of bedrooms, ranging from 0 to 10\

**FinishedBsmtArea (sf)**:\
Format: Numeric (Square Feet)\
Description: Total area of the finished basement in square feet\
Expected Data Range: Sizes of finished basement areas in square feet, ranging from 10 to 6496\

**LotArea(sf)**:\
Format: Numeric (Square Feet)\
Description: Total area of the lot in square feet\
Expected Data Range: Lot areas in square feet, ranging from 0 to 523228\

**AC**:\
Format: Character (logical- boolean)\
Description: Indicates if the property has air conditioning\
Expected Data Range: Yes/No\

**FirePlace**:
Format: Character (logical- boolean)
Description: Indicates if the property has a fireplace\
Expected Data Range: Yes/No\

**Neighborhood**:\
Format: Factor\
Description: Neighborhood area in Ames\
Expected Data Range: Different neighborhood names as factor with 42 levels\



# Question 2: Is there a variable of special interest or focus?


# Question 3: Start the exploration with the main variable

```{r}
# Range 
summary(ames$`Sale Price`)

# NA values 
sum(is.na(ames$`Sale Price`))

# Extreme values 
sum(ames$`Sale Price` > 1000000, na.rm = TRUE)

# Distribution 
library(ggplot2)

ggplot(ames, aes(x = `Sale Price`)) +
  geom_histogram(binwidth = 100000, fill = 'skyblue', color = 'black') +
  labs(title = "Distribution of Sale Prices",
       x = "Sale Price (USD)",
       y = "Frequency")
```

**Main Variable:** Sale Price\
Range: 0 to $20,500,000\
General pattern: Skewed right\ 
Oddities: No NA values but does have a few extreme values and looking at the histogram you can see a few outliers\


# Question 4: Pick a variable that might be related to the main variable Sale Price 

**Variable Chosen:**  Year Built\

```{r}

# Range 
summary(ames$YearBuilt)

# NA values 
sum(is.na(ames$YearBuilt))

# Extreme values 
sum(ames$YearBuilt < 1850, na.rm = TRUE)

# Distribution
a <- na.omit(ames)
ggplot(a, aes(x = YearBuilt)) +
  geom_histogram(binwidth = 5, fill = 'darkgreen', color = 'black') +
  labs(title = "Distribution of Year Built",
       x = "Year Built",
       y = "Frequency")

# Relationship 
ggplot(a, aes(x = YearBuilt, y = `Sale Price`)) +
  geom_point(size = 1.5) +
  labs(title = "Year Built vs Sale Price",
       x = "Year Built",
       y = "Sale Price (USD)")

```

Group member name: Emily Green\
Range: 0 to 2022\
General pattern of variable: Bimodel\ 
Oddities of variable: Lots of NA values and has a extreme value of 0\
Relationship: Looks like a linear relationship but has a few outliers. Since an oddity found for sale price was it having a few outliers it makes sense that this variable has some outliers when looking at their relationship\


Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.
