---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #2

```{r}

library(remotes)
remotes::install_github("heike/classdata")
library(classdata)

```


# Summary of Finding: 
There is a total of 16 variables, with Sale Price being the main variable. It seems there is a positive correlation between Sale Price with Total Living Area and Year Built. There are outliers which we can deduce is a anomaly of luxuriousness. People tend to settle with buying places with 2-4 bedrooms.

# Question 1: Inspect the first few lines of the data set

```{r}

?ames 
colnames(ames)
head(ames)
str(ames)
summary(ames$AC)

```

**Parcel ID**:\
Format: Character\
Description: Unique identifier for each property\
Expected Data Range: Unique IDs, length of characters varies\

**Address**:\
Format: Character\
Description: Property address in Ames, IA\
Expected Data Range: Various addresses in Ames, IA\

**Style**:\
Format: Factor\
Description: Type of housing, such as "Colonial," "Ranch," "Split-Level," etc.\
Expected Data Range: Different housing styles as factor with 12 levels\

**Occupancy**:\
Format: Factor\
Description: Type of occupancy, indicating whether the property is occupied or not\
Expected Data Range: Occupancy status as factor with 5 levels\

**Sale Date**:\
Format: Date\
Description: Date of the property sale\
Expected Data Range: Dates ranging from 2017-07-03 to 2022-08-31\

**Sale Price**:\
Format: Numeric (in US dollars)\
Description: Sales price of the property\
Expected Data Range: Sale prices in US dollars, varying from 0 to 20500000\

**Multi Sale**:\
Format: Character (logical- boolean)\
Description: Indicates if the sale was part of a package or multiple sale\
Expected Data Range: Y/N\

**YearBuilt**:\
Format: Numeric (integer)\
Description: Year in which the house was built\
Expected Data Range: Years ranging from 0 to 2022\

**Acres**:\
Format: Numeric\
Description: Size of the lot in acres\
Expected Data Range: Acres of the lots, ranging from 0 to 12.0120\

**TotalLivingArea (sf)**:\
Format: Numeric (Square Feet)\
Description: Total living area of the property in square feet\
Expected Data Range: Living area sizes in square feet, varying from 0 to 6007\

**Bedrooms**:\
Format: Numeric\
Description: Number of bedrooms in the property\
Expected Data Range: Number of bedrooms, ranging from 0 to 10\

**FinishedBsmtArea (sf)**:\
Format: Numeric (Square Feet)\
Description: Total area of the finished basement in square feet\
Expected Data Range: Sizes of finished basement areas in square feet, ranging from 10 to 6496\

**LotArea(sf)**:\
Format: Numeric (Square Feet)\
Description: Total area of the lot in square feet\
Expected Data Range: Lot areas in square feet, ranging from 0 to 523228\

**AC**:\
Format: Character (logical- boolean)\
Description: Indicates if the property has air conditioning\
Expected Data Range: Yes/No\

**FirePlace**:\
Format: Character (logical- boolean)
Description: Indicates if the property has a fireplace\
Expected Data Range: Yes/No\

**Neighborhood**:\
Format: Factor\
Description: Neighborhood area in Ames\
Expected Data Range: Different neighborhood names as factor with 42 levels\



# Question 2: Is there a variable of special interest or focus?    

Sale Price is the main variable of interest. There are many other variables that are related to it.


# Question 3: Start the exploration with the main variable

```{r}
# Range 
summary(ames$`Sale Price`)

# NA values 
sum(is.na(ames$`Sale Price`))

# Extreme values 
sum(ames$`Sale Price` > 1000000, na.rm = TRUE)

# Distribution 
library(ggplot2)

ggplot(ames, aes(x = `Sale Price`)) +
  geom_histogram(binwidth = 100000, fill = 'skyblue', color = 'black') +
  labs(title = "Distribution of Sale Prices",
       x = "Sale Price (USD)",
       y = "Frequency")
```

**Main Variable:** Sale Price\
Range: 0 to $20,500,000\
General pattern: Skewed right\ 
Oddities: No NA values but does have a few extreme values and looking at the histogram you can see a few outliers\


# Question 4: Pick a variable that might be related to the main variable Sale Price 

**Variable Chosen:**  Year Built\ TotalLivingArea\

```{r}

# Range 
summary(ames$YearBuilt)

# NA values 
sum(is.na(ames$YearBuilt))

# Extreme values 
sum(ames$YearBuilt < 1850, na.rm = TRUE)

# Distribution
a <- na.omit(ames)
ggplot(a, aes(x = YearBuilt)) +
  geom_histogram(binwidth = 5, fill = 'darkgreen', color = 'black') +
  labs(title = "Distribution of Year Built",
       x = "Year Built",
       y = "Frequency")

# Relationship 
ggplot(a, aes(x = YearBuilt, y = `Sale Price`)) +
  geom_point(size = 1.5) +
  labs(title = "Year Built vs Sale Price",
       x = "Year Built",
       y = "Sale Price (USD)")

```    
```{r}
# Range 
summary(ames$`TotalLivingArea (sf)`)

sum(is.na(ames$`TotalLivingArea (sf)`))

summary(ames$`TotalLivingArea (sf)` < 500, na.rm = TRUE)

# Distribution of Total Living Area
b <- na.omit(ames) 
ggplot(b, aes(x = `TotalLivingArea (sf)`)) +
  geom_histogram(binwidth = 100, fill = 'blue', color = 'black') +
  labs(title = "Distribution of Total Living Area",
       x = "Total Living Area (sf)",
       y = "Frequency")

# Relationship with Sale Price
ggplot(a, aes(x = `TotalLivingArea (sf)`, y = `Sale Price`)) +
  geom_point(size = 1.5) +
  labs(title = "Year Built vs Sale Price",
       x = "Total Living Area",
       y = "Sale Price (USD)")
```
Group member name: Emily Green\ Manjul Balayar\
Range: 0 to 2022\
General pattern of variable: Bimodel\ 
Oddities of variable: Lots of NA values and has a extreme value of 0\
Relationship: Looks like a linear relationship but has a few outliers. Since an oddity found for sale price was it having a few outliers it makes sense that this variable has some outliers when looking at their relationship\

``` {r}

summary(ames$`Bedrooms`)

ggplot(ames, aes(x = `Bedrooms`)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Bedrooms",
       x = "Bedrooms",
       y = "Frequency")

ggplot(ames, aes(x = `Bedrooms`, y = `Sale Price`)) +
  geom_point(alpha =0.5, color = "blue")
  labs(title = "Sale Price, vs # of Bedrooms", 
       x = "Bedrooms", 
       y = "Sale Price")

ggplot(ames, aes(x = cut(`Bedrooms`, breaks = 10), y = `Sale Price`)) +
  geom_boxplot(fill = "skyblue", color = "black") +
  labs(title = "Sale Price vs # of Bedrooms ",
       x = "Bedrooms",
       y = "Sale Price")

  
```
Mason Sackpraseuth:
Range: 0-10
Pattern: Histogram is slightly right skewed with some alternate peaks.
Oddities: lots of missing values (N/A). A lot of people tend to go with 3 bedrooms, followed by 4 and then 2. Surprisingly enough there is no sales with 9 bedrooms but there is with 10.
Relationship: There is a non-linear relationship between the number of bedrooms and the sale price. Most of the price seems to clump together except for a few outliers. We can assume these outliers as luxurious homes. There are a number of bedrooms ranging from the 2-4 range.

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.
